# Appendix {#appendix}

## 4D Framework

  This section provides a detailed set of considerations and framework for structuring analytics projects.

### Discover

  During the Discover phase, it is important to remain in the problem zone; seek to understand your clients' needs through active listening and questions. This is not the time for solutioning or committing to any specific deliverables or timelines. If the client's needs are ambiguous, proceeding without clarity is unlikely to result in a favorable outcome.
  
  It is generally helpful to think about analytics solutions -- whether a dashboard with basic metrics and trends or an advanced analysis -- like a Product Owner thinks about the initial and subsequent releases of a commercial product. A **Minimum Viable Product (MVP)** is a version of the solution with the minimum number of features to be useful to early customers who can provide feedback for future enhancements. It is important to clarify that the MVP version of solutions often has both a limited number of users and features, which protects against the tendency to boil the ocean by striving to address every question for every stakeholder. Breaking down large projects into small sets of features that are easier to communicate and adopt provides space for agility and real-time adjustments to the product roadmap per user feedback.
  
  There are some initial questions and considerations that will help frame an analytics project and support its success.
  
  **Client**

  Who is the client? A client can be a person or organization contracting you for consulting services, or an internal stakeholder within your organization who has need. What is important to them?

  **Primary Objective**

  + What is the client ultimately hoping to accomplish? 
  + Is the request merely to satisfy one’s curiosity, or are there actions that can realistically be taken to materially influence the stated objective?

  **Problem Statement**

  + One of the most important early steps is clearly defining the problem statement. If your understanding of the problem -- after translating from the business anecdotes in which it was likely initially expressed -- is misaligned with the client's needs, none of the subsequent steps matter.

  **Guiding Theories**

  + What theoretical explanations can the client offer as potential rationalizations for the phenomena of interest?
  + Are there existing theories in the organizational literature that should guide how the problem is tackled (e.g., findings from similar research implemented in other contexts)?

  **Research Questions**

  To respect the nuances of the problem statement, it is important to unpack it and frame as a set of overarching questions to guide the research.

  + Q1: ...
  + Q2: ...
  + Q3: ...

  To ensure it is possible (and appropriate) to take action in response to these questions, consider the following after drafting each question: "What would I do if I knew the answer?"

  **Research Hypotheses**

  Once research questions are developed, what do you expect to find based on anecdotal stories or empirical findings? As a next step, these expectations should be expressed in the form of research hypotheses. Note that these research hypotheses are different from statistical hypotheses; both will be covered in detail in later chapters.

  + H1: ...
  + H2: ...
  + H3: ...

  To ensure the hypotheses lend to actionable analyses, it is important to consider the following: "What does success look like?" In other words, once the project is complete, against which success measures will the project's success be determined? Curiosity is not a business reason and hope is not a reasonable strategy. The following questions may prove helpful in the promotion of actionable -- over merely interesting -- outcomes:

  + What will be done if the hypotheses are supported?
  + What will be done if the hypotheses are *not* supported?

  **Assumptions**

  At this point, it's helpful to consider what assumptions may be embedded in this discovery work. Are the questions and hypotheses rooted in what the client has theorized, or are these the product of an ambiguous understanding of the client's needs?

  **Cadence**
 
  + Is this analysis a one-off, or could there be a need to refresh this analysis on a regular cadence?
  + Are there dates associated with programs or actions this analysis is intended to support?

  **Aggregation**

  Is there a need for individual-level detail supporting the analysis? Aggregate data should generally be the default unless a compelling justification exists and exception granted from legal and privacy partners. One important role of analysts is to help keep the audience focused on the bigger picture and findings. Access to individual-level detail can not only introduce unnecessary legal and compliance risk but can also lead to questions and probing that can delay taking necessary actions based on the results.

  **Deliverable**

  What is the preferred method of communicating the results of the analysis (e.g., interactive dashboard, static slide deck, document)? It is important to determine this early so that subsequent efforts can be structured to support the preferred deliverable. For example, if an interactive dashboard is preferred, does your Engineering department need to prioritize dependent tasks such as data pipelines, row-level security, BI development, and migrations to production servers?

  **Filters & Dimensions**

  How does your client prefer to segment the workforce? Some common grouping dimensions are business unit, division, team, job family, location, tenure, and management level, but the client may have custom segmentation requirements that will be important to identify and define early in the project.
  
### Design

  Perhaps the most important initial question to answer in the design phase is: "Does anything already exist that addresses part, or all, of the client’s objectives?" If an existing solution will suffice, or a previous analysis can be easily refreshed with recent data, it may be possible to allocate time and resources elsewhere. If related or complimentary analyses have already been performed, they may accelerate new analyses.
  
  The end-user experience is of paramount importance during the Design phase, as solutions should have a consistent look and feel regardless of who developed them. Defining and implementing design guidelines, such as those outlined in Chapter \@ref(data-viz), will ensure consistency across analytics projects as well as within large projects in which multiple analysts are collaborating on various elements of the solution.

  **Data Privacy**

  Are there potential concerns with the study’s objective, planned actions, and/or requested data elements from an employee privacy or legal perspective? A cross-functional data governance committee can help with efficient and consistent decisioning on requests for people data and analytics.
  
  In cases where sensitive attributes such as gender, ethnicity, age, sexual orientation, and disability status are requested, it is best to consult with legal and privacy partners to ensure they green light the intended use of the data. The decision on whether or not to include these sensitive data elements is often less about what the audience can view (e.g., People Partners may already have access to the information at the individual level in the source system) and more anchored in what they plan to do with the information.

  **Data Sources & Elements**
  
  Is the required data already accessible in a data warehouse or other analytics environment? If not, does it need to be? What is required to achieve this?

  + What data sources are required?
  + What data elements are required?

  **Data Quality**

  It is important to understand the data generative process and never make assumptions about how anomalies or missing data should be interpreted. After identifying what data sources will be required for a particular analysis, it is important to meet with source system owners and data stewards to deeply understand the business processes by which data are generated in the system(s). An ideal should always be *unimpeachable* data quality at the analysis presentation stage, and this begins with an early understanding and investigation of data quality in the source systems. Below are some helpful questions to consider:
  
  * Is there a better data source, such as data generated further upstream?
  * Are the data actively validated with automated data quality checks and regular exceptions reports?
  * What is the SLA for the data refreshes, and how often is this SLA met?
  * Is there a clear (and accessible) data source owner/steward?
  * Is there comprehensive documentation and field definitions?
  * Do the tables drive business critical deliverables?
  * Who are the other consumers of the data?

  **Variables**

  How will the constructs be measured (e.g., survey instrument, derived attribute, calculated field)?

  **Analysis Method**

  What are the appropriate analysis methods based on the research hypotheses? If modeling is required, is it more important to index on accuracy or interpretability?

  **Dependencies**

  Are other teams required to develop this solution? What is the nature of the work each dependent team will perform? Are there required system configuration changes? Do these teams have capacity to support?

  **Change Management**

  Will this solution impact current processes or solutions? If so, what is the change management plan to facilitate a seamless transition and user experience?

  **Sign-Off**

  Generally, it is best for the client to signoff on the problem statement, analysis approach, and wire frame for the deliverable (if applicable) before providing an ETA and proceeding to the development phase. This ensures alignment on the client's needs and the perceived utility of the solution in addressing those needs.

### Develop

  While development patterns can vary widely across analytics teams, establishing a set of standards can pay dividends in the form of greater efficiency and reliability over time. Pattern-based development ensures analysts who were not involved in a particular project can access the code and easily and quickly understand each step of the analysis.
  
  **Development Patterns**

  + Are there development patterns that should guide the development approach to support consistency? 
  + Are there existing calculated fields that can/should be leveraged for derived data?
  + Are there best practices that should be employed to optimize performance (e.g., load time for dashboards, executing complex queries during non-peak times)?
  + Are there standard color palettes that should be applied?

  **Productionalizable Code**

  + How do models and data science pipelines need to be developed to facilitate a seamless migration from lower to upper environments? For example, initial EDA may be performed using curated data in flat files for the purpose of identifying meaningful trends, relationships, and differences, but where will this data need to be sourced in production to automate the refresh of models at a regular interval? If the data were provided from multiple source systems, what joins are required to integrate the data? What transformation logic or business rules need to be applied to reproduce the curated data?

  **Unit Testing**

  + What test cases will ensure the veracity of data?
  + Who will perform the testing?

  **User Acceptance Testing (UAT)**

  + In the spirit of agility and constant contact with the client to prevent surprises, it is generally a good idea to have the client take the solution for a test run within the UAT environment and then provide sign-off before migrating to production. If the deliverable is a deck or doc with results from a model, UAT may surface clarifying questions that can be addressed before releasing to the broader audience.

### Deliver

  The Deliver phase can take many forms depending on the solution being released. If the solution is designed for a large user base, a series of recorded trainings may be in order so that there is a helpful reference for those unable to attend the live sessions or new joiners in the future. It is important to monitor success measures, which could be insights aligned to research hypotheses, dashboard utilization metrics, progress following data-informed interventions, or any number of others defined within the Discover phase.

## Data Visualization

  This section provides the code needed to reproduce visuals in Chapter \@ref(data-viz).

### Step-by-Step Visual Upgrade

```{r, eval = FALSE}

# Load libraries
library(dplyr)
library(ggplot2)

# Read employee data
employees <- read.csv("https://raw.githubusercontent.com/crstarbuck/peopleanalytics_book/master/data/employees.csv")

# Create data cube for active employees
smmry_ed_field <- employees %>%
                  dplyr::filter(active == 'Yes') %>%
                  dplyr::count(ed_field)

# Step 1: Build Bar Chart with Defaults
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") 

# Step 2: Remove Legend
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::theme(legend.position = "none")

# Step 3: Assign Colors Strategically
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none")

# Step 4: Add Axis Titles and Margins
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::labs(x = 'Education Field', y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0)),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0)))

# Step 5: Add Left-Justified Title
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::labs(title = 'Life Sciences is the most common education field \npursued by active employees.', x = 'Education Field', y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               plot.title = ggplot2::element_text(hjust = 0),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0)),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0)))

# Step 6: Remove Background
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::labs(title = 'Life Sciences is the most common education field \npursued by active employees.', x = 'Education Field', y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0)),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0)))

# Step 7: Remove Axis Ticks
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::labs(title = 'Life Sciences is the most common education field \npursued by active employees.', x = 'Education Field', y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0)),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0)),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

# Step 8: Mute Titles
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::labs(title = 'Life Sciences is the most common education field \npursued by active employees.', x = 'Education Field', y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

# Step 9: Flip Axes
ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::coord_flip() +
ggplot2::labs(title = 'Life Sciences is the most common education field \npursued by active employees.', x = 'Education Field', y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

# Step 10: Sort Data
ggplot2::ggplot(smmry_ed_field, aes(x = reorder(ed_field, n), y = n, fill = as.factor(ed_field))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::coord_flip() +
ggplot2::labs(title = 'Life Sciences is the most common education field \npursued by active employees.', x = 'Education Field', y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                      "Life Sciences" = "#0070C0", 
                                      "Marketing" = "#BFBFBF", 
                                      "Medical" = "#BFBFBF",
                                      "Other" = "#BFBFBF",
                                      "Technical Degree" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

# Pre/post viz comparison
viz_pre <- ggplot2::ggplot(smmry_ed_field, aes(x = ed_field, y = n, fill = as.factor(ed_field))) +
           ggplot2::geom_bar(stat = "identity") 

viz_post <- ggplot2::ggplot(smmry_ed_field, aes(x = reorder(ed_field, n), y = n, fill = as.factor(ed_field))) +
            ggplot2::geom_bar(stat = "identity") +
            ggplot2::coord_flip() +
            ggplot2::labs(title = 'Life Sciences is the most common education field \npursued by active employees.', x = 'Education Field', y = 'Headcount') +
            ggplot2::scale_fill_manual(values = c("Human Resources" = "#BFBFBF", 
                                                  "Life Sciences" = "#0070C0", 
                                                  "Marketing" = "#BFBFBF", 
                                                  "Medical" = "#BFBFBF",
                                                  "Other" = "#BFBFBF",
                                                  "Technical Degree" = "#BFBFBF")) +
              ggplot2::theme(legend.position = "none",
                             panel.background = ggplot2::element_blank(),
                             axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
                             axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
                             plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
                             axis.ticks.x = ggplot2::element_blank(),
                             axis.ticks.y = ggplot2::element_blank())
  
# Display pre/post visualizations side-by-side
ggpubr::ggarrange(viz_post, viz_pre, ncol = 1, nrow = 2)

```

### Tables

```{r, eval = FALSE}

### Data Prep ###

# Load library
library(dplyr)

# Read employee data
employees <- read.csv("https://raw.githubusercontent.com/crstarbuck/peopleanalytics_book/master/data/employees.csv")

# Append new tenure band column
employees$tenure_band <- dplyr::case_when(
  employees$org_tenure < 1 ~ "Under 1 Year",
  employees$org_tenure < 2.5 ~ "1-2 Years",
  employees$org_tenure < 5.5 ~ "3-5 Years",
  employees$org_tenure <= 10 ~ "6-10 Years",
  TRUE ~ "Over 10 Years"
)

# Store aggregate measures to cube
dept_tenure <- employees %>%
               dplyr::filter(active == 'Yes') %>%
               dplyr::group_by(dept, tenure_band) %>%
               dplyr::summarise(cnt = dplyr::n())

# Specify ordered factor
dept_tenure$tenure_band <- ordered(dept_tenure$tenure_band, levels = c("Under 1 Year", "1-2 Years", "3-5 Years", "6-10 Years", "Over 10 Years"))

# Assign proper case field names
dept_tenure_proper <- dept_tenure %>%
                      rename('Department' = dept,
                             'Tenure' = tenure_band,
                             'Count' = cnt)

```

```{r, eval = FALSE}

### Data Table ###

# Load library
library(DT)

# Format output with datatable
DT::datatable(dept_tenure_proper, rownames = FALSE)

```  

### Heatmaps

```{r, eval = FALSE}

### Data Prep ###

# Load library
library(dplyr)

# Append new tenure band column
employees$tenure_band <- dplyr::case_when(
  employees$org_tenure < 1 ~ "Under 1 Year",
  employees$org_tenure < 2.5 ~ "1-2 Years",
  employees$org_tenure < 5.5 ~ "3-5 Years",
  employees$org_tenure <= 10 ~ "6-10 Years",
  TRUE ~ "Over 10 Years"
)

# Store aggregate measures to cube
dept_tenure <- employees %>%
               dplyr::filter(active == 'Yes') %>%
               dplyr::group_by(dept, tenure_band) %>%
               dplyr::summarise(cnt = dplyr::n())

# Specify ordered factor
dept_tenure$tenure_band <- ordered(dept_tenure$tenure_band, levels = c("Under 1 Year", "1-2 Years", "3-5 Years", "6-10 Years", "Over 10 Years"))

```

```{r, eval = FALSE}

### Heatmap ###

# Load library
library(ggplot2)

ggplot2::ggplot(dept_tenure, aes(x = tenure_band, y = dept, fill = cnt)) +
ggplot2::geom_tile() +
ggplot2::labs(title = 'A large portion of our workforce has between 3 and 10 years\n of company tenure and works within R&D.', x = 'Tenure', y = 'Department') +
ggplot2::guides(fill = guide_legend(title = "Employee Count")) +
ggplot2::scale_fill_continuous(low = "#F2F2F2", high = "#0070C0") +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank(),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"))

```

### Scatterplots

```{r, eval = FALSE}

### Data Prep ###

# Subset df to active sales employees
sales <- subset(employees, dept == 'Sales' & active == 'Yes')

# Set quote attainment flag for data viz coloring
sales$quota_flg <- ifelse(sales$ytd_sales >= 150000, 1, 0)

```

```{r, eval = FALSE}

### Scatterplots ###

# Load libraries
library(ggplot2)
require(scales)

# Basic scatterplot
ggplot2::ggplot(sales, aes(x = work_exp, y = ytd_sales, group = quota_flg)) +
ggplot2::geom_point() +
ggplot2::labs(title = 'The relationship between work experience and YTD sales is positive.', x = 'Work Experience', y = 'YTD Sales (USD)') +
ggplot2::scale_y_continuous(labels = scales::comma) +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               legend.position = "none",
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank(),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"))

# Scatterplot with preattentive attributes
ggplot2::ggplot(sales, aes(x = work_exp, y = ytd_sales, group = quota_flg)) +
ggplot2::geom_point(aes(color = as.factor(quota_flg))) +
ggplot2::geom_hline(yintercept = 150000, linetype = 'dotted', color = "#0070C0") + 
ggplot2::annotate(geom = "text", x = 2, y = 159000, label = "Sales Quota = $150k", color = "#0070C0", size = 3) +
ggplot2::labs(title = 'Some employees with more work experience have \nalready met the full year sales quota.', x = 'Work Experience', y = 'YTD Sales (USD)') +
ggplot2::scale_color_manual(values = c("0" = "#BFBFBF",
                                       "1" = "#0070C0")) +
ggplot2::scale_y_continuous(labels = scales::comma) +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               legend.position = "none",
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank(),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"))

```

### Line Charts

```{r, eval = FALSE}

### Data Prep ###

# Set seed for reproducibility
set.seed(1234)

# Create data
months = 1:24
eng_rt = 5 - runif(1, 2.7, 2.9) + 2.41*months - .41*months^2 + .02*months^3
fin_rt = runif(1, 5, 8) - 6.97 + 15*months - .53*months^2
ppl_rt = 3 - runif(1, 5, 8) - 6.97 + 12*months - .4*months^2
prd_rt = runif(1, 5, 8) - 6.97 + 13*months - .53*months^2

# Combine dimensions and metrics within df
attrition_dat <- data.frame(month = rep(months, 4),
                            dept = c(rep('Engineering', length(months)), 
                                     rep('Finance', length(months)),
                                     rep('People', length(months)),
                                     rep('Product', length(months))),
                            rate = c(eng_rt,
                                     fin_rt,
                                     ppl_rt,
                                     prd_rt))

# Ordered factor for subsequent layering of lines in graph (last value takes the top position)
attrition_dat$dept <- factor(attrition_dat$dept, c("Product", "Finance", "People", "Engineering"))

```

```{r, eval = FALSE}

### Single Series Line Graphs ###

# Load library
library(ggplot2)

ggplot2::ggplot(subset(attrition_dat, dept == 'Engineering'), aes(x = month, y = rate, group = dept)) + 
ggplot2::labs(title = 'Engineering attrition has increased over the last 24 months, \nlargely due to an exponential spike in the last year.', x = 'Month', y = 'Attrition Rate') +
ggplot2::geom_line(color = '#0070C0', size = 1) +
ggplot2::scale_x_continuous(limits = c(0, 25), breaks = seq(0, 24, by = 3)) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```
  
```{r, eval = FALSE}

### Two Series Line Graphs ###

# Load library
library(ggplot2)

ggplot2::ggplot(subset(attrition_dat, dept %in% c('Engineering', 'Product')), aes(x = month, y = rate, color = dept)) + 
ggplot2::labs(title = 'The attrition trends for Engineering and Product \nare negatively correlated over the last year.',
              x = 'Month',
              y = 'Attrition Rate',
              col = 'Department') +
ggplot2::geom_line(size = 1) +
ggplot2::scale_x_continuous(limits = c(0, 27), breaks = seq(0, 24, by = 3)) +
ggplot2::annotate(geom = "text", x = max(attrition_dat$month) + .5, y = attrition_dat[attrition_dat$dept == 'Engineering' & attrition_dat$month == max(attrition_dat$month), 'rate'], label = 'Engineering', color = "#0070C0", size = 4, hjust = 0) +
ggplot2::annotate(geom = "text", x = max(attrition_dat$month) + .5, y = attrition_dat[attrition_dat$dept == 'Product' & attrition_dat$month == max(attrition_dat$month), 'rate'], label = 'Product', color = "#808080", size = 4, hjust = 0) +
ggplot2::scale_color_manual(values = c("Engineering" = "#0070C0",
                                       "Product" = "#BFBFBF")) +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               legend.position = "none",
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```
  
```{r, eval = FALSE}

### Multiple Series Line Graphs ###

# Load library
library(ggplot2)

ggplot2::ggplot(attrition_dat, aes(x = month, y = rate, color = dept)) + 
ggplot2::labs(title = 'The attrition trend for Engineering is quite different \nrelative to attrition patterns for other departments.',
              x = 'Month',
              y = 'Attrition Rate',
              col = 'Department') +
ggplot2::geom_line(size = 1) +
ggplot2::scale_x_continuous(limits = c(0, 27), breaks = seq(0, 24, by = 3)) +
ggplot2::annotate(geom = "text", x = max(attrition_dat$month) + .5, y = attrition_dat[attrition_dat$dept == 'Engineering' & attrition_dat$month == max(attrition_dat$month), 'rate'], label = 'Engineering', color = "#0070C0", size = 4, hjust = 0) +
ggplot2::annotate(geom = "text", x = max(attrition_dat$month) + .5, y = attrition_dat[attrition_dat$dept == 'Product' & attrition_dat$month == max(attrition_dat$month), 'rate'], label = 'Product', color = "#808080", size = 4, hjust = 0) +
ggplot2::annotate(geom = "text", x = max(attrition_dat$month) + .5, y = attrition_dat[attrition_dat$dept == 'Finance' & attrition_dat$month == max(attrition_dat$month), 'rate'], label = 'Finance', color = "#808080", size = 4, hjust = 0) +
ggplot2::annotate(geom = "text", x = max(attrition_dat$month) + .5, y = attrition_dat[attrition_dat$dept == 'People' & attrition_dat$month == max(attrition_dat$month), 'rate'], label = 'People', color = "#808080", size = 4, hjust = 0) +
ggplot2::scale_color_manual(values = c("Engineering" = "#0070C0",
                                       "Product" = "#BFBFBF",
                                       "Finance" = "#BFBFBF",
                                       "People" = "#BFBFBF")) +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               legend.position = "none",
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, margin = ggplot2::margin(t = 0, r = 0, b = 20, l = 0), colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```

### Slopegraphs

```{r, eval = FALSE}

### Data Prep ###

# Build data frame with YoY headcount metrics by department
prepost_scores <- data.frame(date = c(rep('Time 1', 2), rep('Time 2', 2)),
                             group = rep(c('Treatment', 'Control'), 2),
                             score = c(10, 13, 55, 16))

```  

```{r, eval = FALSE}

### Slopegraphs ###

# Load library
library(ggplot2)

# Calculate treatment and control group changes for inclusion in annotation
treatment_chg <- subset(prepost_scores, group == 'Treatment' & date == 'Time 2', select = score) - subset(prepost_scores, group == 'Treatment' & date == 'Time 1', select = score)
control_chg <- subset(prepost_scores, group == 'Control' & date == 'Time 2', select = score) - subset(prepost_scores, group == 'Control' & date == 'Time 1', select = score)

ggplot2::ggplot(prepost_scores, aes(x = date, y = score, group = group)) +
ggplot2::geom_line(aes(color = group), size = 2) +
ggplot2::geom_point(aes(color = group), size = 4) +
ggplot2::expand_limits(y = 0) +
ggplot2::annotate(geom = "text", x = 2.1, y = prepost_scores[prepost_scores$group == 'Treatment' & prepost_scores$date == 'Time 2', 'score'] - 2, label = paste0(prepost_scores[prepost_scores$group == 'Treatment' & prepost_scores$date == 'Time 2', 'score'], "% (", ifelse(treatment_chg >= 0, '+', ''), treatment_chg, ")", "\nTreatment Group"), color = "#0070C0", size = 4, hjust = 0) +
ggplot2::annotate(geom = "text", x = 2.1, y = prepost_scores[prepost_scores$group == 'Control' & prepost_scores$date == 'Time 2', 'score'] - 2, label = paste0(prepost_scores[prepost_scores$group == 'Control' & prepost_scores$date == 'Time 2', 'score'], "% (", ifelse(control_chg >= 0, '+', ''), control_chg, ")", "\nControl Group"), color = "#808080", size = 4, hjust = 0) +
ggplot2::labs(title = '          A more favorable change in engagement was observed \n          for the treatment group relative to \n          the control group.', x = 'Observation Period') +
ggplot2::scale_color_manual(values = c("Control" = "#BFBFBF",
                                       "Treatment" = "#0070C0")) +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               legend.position = "none",
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank(),
               axis.text.y = ggplot2::element_blank(),
               plot.title = ggplot2::element_text(colour = "#404040"),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_blank())

```

### Bar Charts
  
```{r, eval = FALSE}

### Data Prep ###

# Set seed for reproducibility
set.seed(1234)

# Generate favorability distributions
fav_pct <- round(runif(7, 20, 45), 0)
neu_pct <- round(runif(7, 20, 45), 0)
unfav_pct <- 100 - fav_pct - neu_pct
scores <- c(fav_pct, neu_pct, unfav_pct)

# Average top box (favorable) score
topbox_avg <- round(mean(fav_pct), 0)

# Build data frame with YoY headcount metrics by department
engagement_scores <- data.frame(dept = rep(c('Engineering', 'Finance', 'Legal', 'Marketing', 'People', 'Product', 'Sales'), 3),
                                favorability = c(rep('Favorable', 7), rep('Neutral', 7), rep('Unfavorable', 7)),
                                pct = scores)

# Rank departments by top box score to support sorting
dept_rank <- engagement_scores %>%
             dplyr::filter(favorability == 'Favorable') %>% 
             dplyr::arrange(desc(pct)) %>% 
             dplyr::mutate(rank = dense_rank(desc(pct))) %>%
             dplyr::select(dept, rank)

# Flag top department records in df to support conditional coloring
engagement_scores$top_score = ifelse(engagement_scores$dept == dept_rank[1, 'dept'], 1, 0)

# Add department rank based on top box scores
engagement_scores <- left_join(engagement_scores, dept_rank, by = "dept")

```  
  
```{r, eval = FALSE}

### Vertical Bar Charts ###

ggplot2::ggplot(subset(engagement_scores, favorability == 'Favorable'), aes(x = reorder(dept, -pct), y = pct, fill = as.factor(top_score))) +
ggplot2::labs(title = 'The People team is leading in department-level engagement.', x = 'Department', y = 'Engagement Score') +
ggplot2::geom_bar(stat = "identity") +
ggplot2::geom_hline(yintercept = topbox_avg, linetype = 'dotted', color = "#404040") + 
ggplot2::annotate(geom = "text", x = 7.1, y = topbox_avg + 1.3, label = paste0("Avg = ", topbox_avg), color = "#404040", size = 3) +
ggplot2::scale_fill_manual(values = c("1" = "#0070C0",
                                      "0" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 0, b = 20, l = 0), colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```

```{r, eval = FALSE}

### Horizontal Bar Charts ###

ggplot2::ggplot(subset(engagement_scores, favorability == 'Favorable'), aes(x = reorder(dept, pct), y = pct, fill = as.factor(top_score))) +
ggplot2::labs(title = 'The People team is leading in department-level engagement.', x = 'Department', y = 'Engagement Score') +
ggplot2::geom_bar(stat = "identity") +
ggplot2::coord_flip() +
ggplot2::geom_hline(yintercept = topbox_avg, linetype = 'dotted', color = "#404040") + 
ggplot2::annotate(geom = "text", x = 1, y = topbox_avg + 2.3, label = paste0("Avg = ", topbox_avg), color = "#404040", size = 3) +
ggplot2::scale_fill_manual(values = c("1" = "#0070C0",
                                      "0" = "#BFBFBF")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 0, b = 20, l = 0), colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```
  
```{r, eval = FALSE}

### Stacked Bar Charts ###

ggplot2::ggplot(engagement_scores, aes(x = reorder(dept, -rank), y = pct, fill = factor(favorability, levels = c("Unfavorable", "Neutral", "Favorable")))) +
ggplot2::labs(title = 'Half of Engineering respondents report unfavorable engagement.', x = 'Department', y = 'Engagement Score') +
ggplot2::guides(fill = guide_legend(title = "Favorability")) +
ggplot2::geom_bar(position = "fill", stat = "identity") +
ggplot2::coord_flip() +
ggplot2::scale_fill_manual(values = c("Favorable" = "#0070C0",
                                      "Neutral" = "#D9D9D9",
                                      "Unfavorable" = "#ED7D31")) +
ggplot2::scale_y_continuous(labels = scales::percent) +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 0, b = 20, l = 0), colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```

### Combination Charts

```{r, eval = FALSE}

### Data Prep ###

# Set seed for reproducibility
set.seed(12345)

# Turnover rates
reg_rt <- round(runif(24, 5, 15), 1)
nonreg_rt <- round(runif(24, 1, 5), 1)
vol_rt <- reg_rt + nonreg_rt
turnover_dat <- data.frame(month = 1:24,
                           reg_rt = reg_rt,
                           nonreg_rt = nonreg_rt,
                           vol_rt = vol_rt)

```

```{r, eval = FALSE}

### Combination Charts ###

# Load library
library(ggplot2)

ggplot2::ggplot(turnover_dat) + 
ggplot2::labs(title = 'The largest share of monthly voluntary attrition over the 24-month period \nhas consistently been regrettable terminations.',
              x = 'Month',
              y = 'Attrition Rate') +
ggplot2::geom_bar(aes(x = month, y = vol_rt), fill = '#E7E6E6', stat = "identity") +
ggplot2::geom_line(aes(x = month, y = reg_rt), size = 1, color = "#ED7D31", group = 1) +
ggplot2::geom_line(aes(x = month, y = nonreg_rt), size = 1, color = "#9BC2E6", group = 1) +
ggplot2::scale_x_continuous(limits = c(0, 27), breaks = seq(0, 24, by = 3)) +
ggplot2::scale_y_continuous(limits = c(0, 20), breaks = seq(0, 20, by = 5)) +
ggplot2::annotate(geom = "text", x = max(turnover_dat$month) + .7, y = turnover_dat[turnover_dat$month == max(turnover_dat$month), 'reg_rt'], label = 'Regrettable', color = "#ED7D31", size = 4, hjust = 0) +
ggplot2::annotate(geom = "text", x = max(turnover_dat$month) + .7, y = turnover_dat[turnover_dat$month == max(turnover_dat$month), 'nonreg_rt'], label = 'Not\nRegrettable', color = "#5B9BD5", size = 4, hjust = 0) +
ggplot2::theme(panel.background = ggplot2::element_blank(),
               legend.position = "none",
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```

### Waterfall Charts

```{r, message = FALSE, warning = FALSE}

### Data Prep ###

# Generate headcount data with id field to define bar order
hc_dat <- data.frame(id = 1:6,
                     event = c("Starting HC", "Hires", "Transfers In", "Transfers Out", "Exits", "Ending HC"),
                     type = c("Headcount", "Growth", "Growth", "Loss", "Loss", "Headcount"),
                     count = c(100, 50, 10, -10, -20, 130),
                     start = NA,
                     end = NA)

# Define start and end values to support waterfall chart
hc_dat$end <- cumsum(hc_dat$count)
hc_dat$end <- c(head(hc_dat$end, -1), 0)
hc_dat$start <- c(0, head(hc_dat$end, -1))

# Swap start/end values for last record (Ending HC)
hc_dat[nrow(hc_dat), "end"] <- hc_dat[nrow(hc_dat), "start"]
hc_dat[nrow(hc_dat), "start"] <- 0

```

```{r, eval = FALSE}

### Waterfall Chart ###

# Load library
library(ggplot2)

ggplot2::ggplot(hc_dat, aes(x = event, fill = type)) +
ggplot2::geom_rect(aes(x = reorder(event, id),
                       xmin = id - 0.45,
                       xmax = id + 0.50,
                       ymin = end,
                       ymax = start)) +
geom_text(aes(id, end, label = ifelse(id %in% c(1,6), end, '')), # Only label start and end HC
              vjust = 1.5) +
ggplot2::labs(title = 'Hires and inbound transfers have outpaced outbound transfers and exits \nin Engineering, leading to a net YTD increase in headcount.',
              x = 'Event',
              y = 'Headcount') +
ggplot2::scale_fill_manual(values = c("Headcount" = "#D9D9D9",
                                      "Growth" = "#0070C0",
                                      "Loss" = "#ED7D31")) +
ggplot2::theme(legend.position = "none",
               panel.background = ggplot2::element_blank(),
               axis.title.x = ggplot2::element_text(margin = ggplot2::margin(t = 10, r = 0, b = 0, l = 0), colour = "#404040"),
               axis.title.y = ggplot2::element_text(margin = ggplot2::margin(t = 0, r = 10, b = 0, l = 0), colour = "#404040"),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank())

```

### Waffle Charts

```{r, message = FALSE, warning = FALSE}

### Data Prep ###

# Create df with TA funnel metrics
ta_dat <- data.frame(stage = c("Apply", "Phone Screen", "Interview", "Offer Extend", "Offer Accept"),
                     cnt = c(60, 20, 10, 6, 4))

# Set depth of waffle chart (# of y-axis rows)
depth <- 10

# Each observation needs an x and y coordinate, and y needs to be specified first for a waffle chart with horizontal accumulation
waffle_dat <- expand.grid(y = 1:depth,
                          x = seq_len(ceiling(sum(ta_dat$cnt) / depth)))

# Expand the applicant counts into a vector of stage labels
stages <- rep(ta_dat$stage, ta_dat$cnt)

# Integrate stages and fill any extra tiles with NA
waffle_dat$stage <- c(stages, rep(NA, nrow(waffle_dat) - length(stages)))

```

```{r, eval = FALSE}

### Waffle Chart ###

# Load library
library(ggplot2)

ggplot2::ggplot(waffle_dat, aes(x = x, y = y, fill = stage)) + 
ggplot2::labs(title = "We select 10 in every 100 applicants. \nOur offer acceptance rate (OAR) is 40%.") + 
ggplot2::geom_tile(color = "white") +
ggplot2::scale_fill_manual(values = c("Apply" = "#D9D9D9", 
                                      "Phone Screen" = "#BFBFBF", 
                                      "Interview" = "#A6A6A6", 
                                      "Offer Extend" = "#808080",
                                      "Offer Accept" = "#0070C0")) +
ggplot2::scale_y_reverse() +
ggplot2::theme_void() +
ggplot2::theme(legend.title = ggplot2::element_blank(),
               plot.title = ggplot2::element_text(hjust = 0, colour = "#404040"))

```

### Sankey Diagrams

```{r, eval = FALSE}

### Data Prep ###

# Load library
library(dplyr)

# Set seed for reproducibility
set.seed(1234)

# Create nodes df
nodes <- data.frame(name = c('Engineering', 'Finance', 'Legal', 'Marketing', 'People', 'Product', 'Sales', # source
                             'Engineering', 'Finance', 'Legal', 'Marketing', 'People', 'Product', 'Sales')) # destination

# Create links df
links <- expand.grid(source = 0:6, target = 7:13)

# Append employee transfer counts per department pair to links df
links$value <- ifelse(links$source == links$target-7, round(rnorm(1000, 150, 50), 0), round(rnorm(1000, 30, 5), 0))

# Append source department variable for colored connections
links$dept <- dplyr::case_when(
  links$source == 0 ~ "Engineering",
  links$source == 1 ~ "Finance",
  links$source == 2 ~ "Legal",
  links$source == 3 ~ "Marketing",
  links$source == 4 ~ "People",
  links$source == 5 ~ "Product",
  links$source == 6 ~ "Sales",
  TRUE ~ "NA"
)

# Append within/outside department transfer variable for colored connections
links$wiout_dept <- dplyr::case_when(
  (links$source == 0 & links$target == 7) | (links$source == 1 & links$target == 8) | (links$source == 2 & links$target == 9) | (links$source == 3 & links$target == 10) | (links$source == 4 & links$target == 11) | (links$source == 5 & links$target == 12) | (links$source == 6 & links$target == 13) ~ "Within",
  TRUE ~ "Outside"
)

# Append dichotomous product/other indicator to links and nodes data frames
links$prod <- ifelse(links$dept == 'Product', 'Product', 'Other')
nodes$prod <- ifelse(nodes$name == 'Product', 'Product', 'Other')

```

```{r, eval = FALSE}

### Sankey Diagrams ###

# Load library
library(networkD3)

# Sankey diagram 1
networkD3::sankeyNetwork(Links = links,
                         Nodes = nodes,
                         Source = "source", 
                         Target = "target",
                         Value = "value", 
                         NodeID = "name",
                         fontSize = 12, 
                         nodeWidth = 30)

# Sankey diagram 2
networkD3::sankeyNetwork(Links = links,
                         Nodes = nodes,
                         Source = "source", 
                         Target = "target",
                         Value = "value", 
                         NodeID = "name",
                         fontSize = 12, 
                         nodeWidth = 30, 
                         LinkGroup = "dept")

# Sankey diagram 3
networkD3::sankeyNetwork(Links = links,
                         Nodes = nodes,
                         Source = "source", 
                         Target = "target",
                         Value = "value", 
                         NodeID = "name",
                         fontSize = 12, 
                         nodeWidth = 30, 
                         LinkGroup = "wiout_dept")

# Sankey diagram 4
# Define color palette
colorJS <- paste('d3.scaleOrdinal(["#BFBFBF", "#0070C0"])')

networkD3::sankeyNetwork(Links = links,
                         Nodes = nodes,
                         Source = "source", 
                         Target = "target",
                         Value = "value", 
                         NodeID = "name",
                         fontSize = 12, 
                         nodeWidth = 30, 
                         NodeGroup = "prod",
                         LinkGroup = "prod",
                         colourScale = colorJS)

```

### Pie Charts

```{r, EVAL = FALSE}

### Data Prep ###

# Create data cube for gender representation among active employees
smmry_gender <- data.frame(gender = c('Male', 'Female'),
                           cnt = c(732, 501),
                           pct = c(59.4, 40.6))

```

```{r, eval = FALSE}

### Pie Chart ###

# Load library
library(ggplot2)

ggplot2::ggplot(smmry_gender, aes(x = "", y = pct, fill = as.factor(gender))) +
ggplot2::geom_bar(stat = "identity") +
ggplot2::coord_polar("y", start = 0) +
ggplot2::geom_text(aes(label = paste0(pct, "%")), color = ifelse(smmry_gender$gender == 'Male', 'white', 'black'), position = position_stack(vjust = 0.5)) +
ggplot2::labs(title = 'Nearly 60 percent of active employees identify as male.') +
ggplot2::scale_fill_manual(values = c("Female" = "#BFBFBF", 
                                      "Male" = "#0070C0")) +
ggplot2::theme_void() +
ggplot2::theme(legend.title = ggplot2::element_blank(),
               panel.background = ggplot2::element_blank(),
               axis.ticks.x = ggplot2::element_blank(),
               axis.ticks.y = ggplot2::element_blank(),
               plot.title = ggplot2::element_text(colour = "#404040"))

```
